---
title: "Fig2_PGP_Analysis"
author: "Shawn Higdon"
date: "2/10/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Libraries
```{r, message=FALSE}
library(treeio)
library(ggtree)
library(ggnewscale)
library(ComplexHeatmap)
library(tidyverse)
library(dendextend)
library(ape)
library(circlize)
library(viridis)
library(randomcoloR)
library(RColorBrewer)
library(cowplot)
```

# Read in Genome Metadata

> ABB to BCW Genome ID Map

> Metabat Bin Count per Isolate

> Sourmash LCA Classification Data

> Genome Assembly Statistics: Quast output and coverage estimation

```{r}
# ABB to BCW ID MAP
abb_bcw_labels <- read.csv("./genome_metadata/abb_genome_bcw_labels.csv", header = T)

# METABAT BIN INFO
metabat_bins <- read_csv("./genome_metadata/metabat_bin_count.csv", col_names = c("ID", "n_bins"))

# Sourmash LCA Genome Classification Data
smash_lca_whole_genomes <- read.csv("./sourmash_data/sm-lca-whole_genomes-all-k31.csv", header = T)

# Read in quast Summary for all isolate assemblies done with MEgahit
asm_stats <- read_tsv("genome_metadata/assembly_stats/all_quast_reports.tsv") # includes eukaryotic isolates

# Read in assembly coverage stat table
asm_cov <- read.table("genome_metadata/assembly_stats/avg_coverage.tsv", header = T)

# Add AVG_COV to asm_stats
asm_stats <- inner_join(asm_stats, asm_cov, by = "Assembly")

```

## Add BCW labels to all metadata files
```{r}
# Add BCW label to metabat bin DF
metabat_bins$bcw_id <- abb_bcw_labels$BCW_ID[match(metabat_bins$ID, abb_bcw_labels$ABB_ID)]

# Add BCW label to sourmash lca data
smash_lca_whole_genomes$bcw_id <- abb_bcw_labels$BCW_ID[match(smash_lca_whole_genomes$ID, abb_bcw_labels$ABB_ID)]

## change '_' to '-' for bcw_id
smash_lca_whole_genomes$bcw_id <- sub("_", "-", smash_lca_whole_genomes$bcw_id)

## empty cells are 'unassigned'
smash_lca_whole_genomes$genus <- sub("^$", "unassigned", smash_lca_whole_genomes$genus)
```

## Generate list of bacterial isolates with single genome bin
```{r}
# create new dataframe
single_isolates <- metabat_bins %>% select( "bcw_id", "ID", "n_bins")
single_isolates$bcw_id <- sub("_", "-", single_isolates$bcw_id)

# add genus data
single_isolates$genus <- smash_lca_whole_genomes$genus[match(single_isolates$bcw_id, smash_lca_whole_genomes$bcw_id)]

# add phylum data
single_isolates$phylum <- smash_lca_whole_genomes$phylum[match(single_isolates$bcw_id, smash_lca_whole_genomes$bcw_id)]

# empty cells for genus are 'unassigned'
single_isolates$genus <- sub("^$", "unassigned", single_isolates$genus)

# empty cells for Phylum are 'unassigned'
single_isolates$phylum <- sub("^$", "unassigned", single_isolates$phylum)

# remove eukaryotes
single_isolates <- single_isolates %>% filter(genus != "Meyerozyma" &
                                                genus != "Rhodotorula")
# remove isolates that were pulled apart
single_isolates <- single_isolates %>% filter(bcw_id != "BCW_201831.1" &
                                                bcw_id != "BCW_201831.2" &
                                                bcw_id != "BCW_200557.1" &
                                                bcw_id != "BCW_200557.2")
# Keep only isolates with 1 genomic bin
single_isolates <- single_isolates %>% filter(n_bins == 1)

# add abb_ids
single_isolates$abb_id <- abb_bcw_labels$ABB_ID[match(single_isolates$bcw_id, abb_bcw_labels$BCW_ID)]

# change '_' to '-' for bcw_id
single_isolates$bcw_id <- sub("_", "-", single_isolates$bcw_id)

```

## BNF Assay Data

> Read in the Data for BNF assay

```{r, message=FALSE}

# Read in the BNF Assay raw data file
pgp_15N <- read_csv("./pgp_data/PGP_15N.csv")

## change '_' to '-' for bcw_id of pgp_15N
pgp_15N$BCW_ID <- sub("_", "-", pgp_15N$BCW_ID)
```

### Sourmash Compare Data

> Read in comparative matrix and associated data from Sourmash comparison of whole genome sequences.

```{r}
# read in sourmash k31 compare matrix for single isolate genomes
sm_k31_cmp <- read.csv("sourmash_data/pure_isolate_492_cmp/pure_isolate_k31_cmp_492.csv", header=T, check.names = F)

# set rownames to colnames
rownames(sm_k31_cmp) <- colnames(sm_k31_cmp)
#head(rownames(sm_k31_cmp))

# add bcw ids to sm_k31_cmp data frame

## change '_' to '-' for BCW ids
abb_bcw_labels$BCW_ID <- sub("_", "-", abb_bcw_labels$BCW_ID)

## match BCW ids
sm_k31_cmp$bcw_id <- abb_bcw_labels$BCW_ID[match(rownames(sm_k31_cmp), abb_bcw_labels$BCW_ID)]

# add genus info to sm_k31
sm_k31_cmp$genus <- single_isolates$genus[match(sm_k31_cmp$bcw_id, single_isolates$bcw_id)]

# Create tx_name Variable with Hybrid genus + Isolate Name: GENUS sp. BCWXXXX
sm_k31_cmp$tx_name <- paste(sm_k31_cmp$genus, sprintf("sp. %s", sm_k31_cmp$bcw_id))

# create matrix
sm_k31_cmp_mat <- as.matrix(sm_k31_cmp[,1:492])

rownames(sm_k31_cmp_mat) <- sm_k31_cmp$tx_name
colnames(sm_k31_cmp_mat) <- sm_k31_cmp$tx_name

```

## PGP Master Dataframe

> Make a data frame with results from:

1. 15N Assay

2. Sourmash Comparisons

### Make the Data Frame
```{r}
# Make pgp master dataframe
pgp_master <- data.frame(TX_ID=sm_k31_cmp$tx_name)

# add BCW ids
pgp_master$BCW_ID <- sm_k31_cmp$bcw_id

# add ABB ids
pgp_master$ABB_ID <- abb_bcw_labels$ABB_ID[match(pgp_master$BCW_ID, abb_bcw_labels$BCW_ID)]

# add N15 ratio values for each isolate
pgp_master$`BNF` <- pgp_15N$ratio[match(pgp_master$BCW_ID, pgp_15N$BCW_ID)]

# add Genus classification for each isolate
pgp_master$Genus <- smash_lca_whole_genomes$genus[match(pgp_master$BCW_ID, smash_lca_whole_genomes$bcw_id)]
pgp_master$Genus <- as.factor(pgp_master$Genus) # coerce to factor

## assign a unique color to each genus
set.seed(8888)
pgp_genus_colors <- data.frame(Genus = unique(pgp_master$Genus), Color =  distinctColorPalette(length(unique(pgp_master$Genus))))

# add phylum classification for each isolate
pgp_master$Phylum <- smash_lca_whole_genomes$phylum[match(pgp_master$BCW_ID, smash_lca_whole_genomes$bcw_id)]

## assign a unique color to each phylum
set.seed(1234)
pgp_phylum_colors <- data.frame(Phylum = unique(pgp_master$Phylum), Color =  distinctColorPalette(length(unique(pgp_master$Phylum))))

### match Genus
pgp_master$Genus_Color <- pgp_genus_colors$Color[match(pgp_master$Genus, pgp_genus_colors$Genus)]

### match Phylum
pgp_master$Phylum_Color <- pgp_phylum_colors$Color[match(pgp_master$Phylum, pgp_phylum_colors$Phylum)]

## change number of significant digits for numeric values to 3
pgp_master <- pgp_master %>% mutate_if(is.numeric, round, digits = 3)

# Remove incomplete cases (Isolates with NA values for bnf assays)
#pgp_master <- drop_na(pgp_master)

# Create list of isolates with complete pgp cases
single_isolates_complete_cases <- data.frame(BCW_ID = pgp_master$BCW_ID)
#write_csv(single_isolates_complete_cases, "./R_output_files/single_isolates_complete_cases.csv", col_names = F)

# add sm_k31_cmp data to pgp_master
pgp_master <- cbind(sm_k31_cmp_mat, pgp_master)

pgp_master %>%
  group_by(Genus ) %>%
  count()

str(pgp_master[,493:500])
```

> Used list of 492 isolates with single bin genomes to perform `sourmash compare`.

```
sourmash compare -k 31
```
### All Single Isolates
```{r}
# Create Phylo Tree from sourmash matrix

sm_k31_tree_mat <- dist2(sm_k31_cmp_mat)

sm_k31_tree_fit <- hclust(sm_k31_tree_mat)

sm_k31_phylo <- as.phylo(sm_k31_tree_fit)

# set rownames of pgp_master to match tree tip labels
rownames(pgp_master) <- sm_k31_phylo$tip.label

head(rownames(pgp_master))

colnames(pgp_master[,493:500])

levels(pgp_master$Genus)
```

#### BNF Phenotype Tree
```{r}
# Make tree
sm_k31_tree_o15 <- ggtree(sm_k31_phylo, color = "black", size = 0.3, layout = "fan", open.angle = 15)

# Genus Tree
bnf0_notip <- gheatmap(sm_k31_tree_o15, pgp_master[, "Genus", drop = F],
                     offset = 0,
                     width = 0.05,
                     colnames_position = "top",
                     colnames_angle = 0,
                     colnames_offset_y = 9.5,
                     font.size = 5.5,
                     color = "grey50") + 
  scale_fill_manual(values=c("#E6BE53",
                             "#CAD773",
                             "#81D870",
                             "#8F3EE4",
                             "#E25F9D",
                             "#D8E9D6",
                             "#72A890",
                             "#CDEC4C",
                             "#C59D6B",
                             "#86D6E1",
                             "#77E94F",
                             "#81748E",
                             "#64E7A9",
                             "#D3D0E3",
                             "#767CDE",
                             "#7DAEDE",
                             "#E2684F",
                             "#E1ABAA",
                             "#7BEFDF",
                             "#DA47D2",
                             "#D7A7DF",
                             "#E5DEA8",
                             "#B4E3A7",
                             "#CF7DDC"),
                    name = "Genus",
                    guide = guide_legend(
                      direction = "vertical",
                      ncol = 2,
                      title.position = "top",
                      title.theme = element_text(size = 18, hjust = 0.5),
                      label.position = "right",
                      label.theme = element_text(size = 14)
                      )) + theme(legend.title = element_text(),
                                 legend.key.size = unit(.5, "cm"))

bnf_p0 <- bnf0_notip + new_scale_fill()

# Add gheatmap 1 for BNF Ratio Data

bnf_p1 <- gheatmap(bnf_p0, pgp_master[, "BNF", drop = F],
                     offset = 0.25,
                     width = 0.05,
                     colnames_position = "top",
                     colnames_angle = 0,
                     colnames_offset_y = 9.5,
                     font.size = 5.5,
                     color = "grey50") + 
     scale_fill_viridis_c(option = "inferno",
                       direction = -1,
                       name = "BNF (15N/14N Ratio)",
                       limits = c(0, 5),
                       breaks = c(0, 1, 2, 3, 4, 5),
                       guide = guide_colorbar(
                         direction = "horizontal",
                         title.position = "top",
                         title.theme = element_text(size = 18, hjust = 0.5),
                         barwidth = 18,
                         label.position = "bottom",
                         label.theme = element_text(size = 14)
                         )
                       ) +
  theme(legend.title = element_text(), legend.position = "right")



bnf_rot <- print(rotate_tree(bnf_p1, -82))

# save plot
ggsave("./Plots/bnf_rot.pdf", bnf_rot, width = 22, height = 14)

```

## NIF Groups

### NIF Group BNF DF

> Make a data frame with results from:

1. 15N Assay

2. Sourmash Comparisons

#### Make the Data Frame
```{r}
# Make pgp master dataframe
pgp_master <- data.frame(TX_ID=sm_k31_cmp$tx_name)

# add BCW ids
pgp_master$BCW_ID <- sm_k31_cmp$bcw_id

# add ABB ids
pgp_master$ABB_ID <- abb_bcw_labels$ABB_ID[match(pgp_master$BCW_ID, abb_bcw_labels$BCW_ID)]

# add N15 ratio values for each isolate
pgp_master$`BNF` <- pgp_15N$ratio[match(pgp_master$BCW_ID, pgp_15N$BCW_ID)]

# add Genus classification for each isolate
pgp_master$Genus <- smash_lca_whole_genomes$genus[match(pgp_master$BCW_ID, smash_lca_whole_genomes$bcw_id)]
pgp_master$Genus <- as.factor(pgp_master$Genus) # coerce to factor

## assign a unique color to each genus
set.seed(8888)
pgp_genus_colors <- data.frame(Genus = unique(pgp_master$Genus), Color =  distinctColorPalette(length(unique(pgp_master$Genus))))

# add phylum classification for each isolate
pgp_master$Phylum <- smash_lca_whole_genomes$phylum[match(pgp_master$BCW_ID, smash_lca_whole_genomes$bcw_id)]

## assign a unique color to each phylum
set.seed(1234)
pgp_phylum_colors <- data.frame(Phylum = unique(pgp_master$Phylum), Color =  distinctColorPalette(length(unique(pgp_master$Phylum))))

### match Genus
pgp_master$Genus_Color <- pgp_genus_colors$Color[match(pgp_master$Genus, pgp_genus_colors$Genus)]

### match Phylum
pgp_master$Phylum_Color <- pgp_phylum_colors$Color[match(pgp_master$Phylum, pgp_phylum_colors$Phylum)]

## change number of significant digits for numeric values to 3
pgp_master <- pgp_master %>% mutate_if(is.numeric, round, digits = 3)

# Remove incomplete cases (Isolates with NA values for bnf assays)
#pgp_master <- drop_na(pgp_master)

# Create list of isolates with complete pgp cases
single_isolates_complete_cases <- data.frame(BCW_ID = pgp_master$BCW_ID)
#write_csv(single_isolates_complete_cases, "./R_output_files/single_isolates_complete_cases.csv", col_names = F)

# add sm_k31_cmp data to pgp_master
pgp_master <- cbind(sm_k31_cmp_mat, pgp_master)

pgp_master %>%
  group_by(Genus ) %>%
  count()

str(pgp_master[,493:500])
```

#### Add DS Group Info to PGP Master
```{r}
# read in dos santos group membership metadata file
 ds_groups <- read_csv("./genome_metadata/ds_groups.csv", col_types = "cf")

# add ds group to each row of pgp master
pgp_master$nif_group <- ds_groups$group[match(pgp_master$BCW_ID, ds_groups$bcw_id)]

```


> Read Sourmash K31 Compare Data for Each NIF Group

### Dos-Santos Positive (DSP) - Group A
```{r}
# DSP Group
smash_dsp_cmp <- read.csv("./sourmash_data/dsp_dos-santos-positive/pure_iso_dsp_k31_cmp.csv", header = TRUE, check.names = FALSE)

# set rownames to colnames
rownames(smash_dsp_cmp) <- colnames(smash_dsp_cmp)

# add bcw ids to smash_dsp_cmp data frame
smash_dsp_cmp$bcw_id <- abb_bcw_labels$BCW_ID[match(rownames(smash_dsp_cmp), abb_bcw_labels$BCW_ID)]

# add genus info to smash_dsp_cmp
smash_dsp_cmp$Genus <- smash_lca_whole_genomes$genus[match(smash_dsp_cmp$bcw_id, smash_lca_whole_genomes$bcw_id)]

# add tx_name Variable to smash_dsp_cmp
smash_dsp_cmp$tx_name <- paste(smash_dsp_cmp$genus, sprintf("sp. %s", smash_dsp_cmp$bcw_id))

# add BNF value
smash_dsp_cmp$BNF <- pgp_15N$ratio[match(smash_dsp_cmp$bcw_id, pgp_15N$BCW_ID)]

# set rownames to match sample
rownames(smash_dsp_cmp) <- smash_dsp_cmp$tx_name

# create matrices for tree

## sourmash compare matrix
smash_dsp_cmp_mat <- as.matrix(smash_dsp_cmp[,1:193])
rownames(smash_dsp_cmp_mat) <- smash_dsp_cmp$tx_name
colnames(smash_dsp_cmp_mat) <- rownames(smash_dsp_cmp_mat)

```

#### DSP tree objects
```{r}
# Create Phylo Tree from sourmash matrix

smash_dsp_k31_tree_mat <- dist2(smash_dsp_cmp_mat)
smash_dsp_k31_tree_fit <- hclust(smash_dsp_k31_tree_mat)
smash_dsp_k31_phylo <- as.phylo(smash_dsp_k31_tree_fit)

```


#### DSP ggtree
```{r}
# Make tree
sm_dsp_k31_tree <- ggtree(smash_dsp_k31_phylo,
                          color = "black",
                          size = 0.4,
                          layout = "fan",
                          open.angle = 15)
```

### Semi Dos-Santos (SDS) - Group B
```{r}
# SDS Group
smash_sds_cmp <- read.csv("./sourmash_data/sds_semi-dos-santos/pure_iso_sds_k31_cmp.csv", header = TRUE, check.names = FALSE)

# set rownames to colnames
rownames(smash_sds_cmp) <- colnames(smash_sds_cmp)

# add bcw ids to smash_sds_cmp data frame
smash_sds_cmp$bcw_id <- abb_bcw_labels$BCW_ID[match(rownames(smash_sds_cmp), abb_bcw_labels$BCW_ID)]

# add genus info to smash_sds_cmp
smash_sds_cmp$Genus <- smash_lca_whole_genomes$genus[match(smash_sds_cmp$bcw_id, smash_lca_whole_genomes$bcw_id)]

# add tx_name Variable to smash_sds_cmp
smash_sds_cmp$tx_name <- paste(smash_sds_cmp$genus, sprintf("sp. %s", smash_sds_cmp$bcw_id))

# add BNF value
smash_sds_cmp$BNF <- pgp_15N$ratio[match(smash_sds_cmp$bcw_id, pgp_15N$BCW_ID)]

# set rownames to match sample
rownames(smash_sds_cmp) <- smash_sds_cmp$tx_name

# create matrices for tree

## sourmash compare matrix
smash_sds_cmp_mat <- as.matrix(smash_sds_cmp[,1:66])
rownames(smash_sds_cmp_mat) <- smash_sds_cmp$tx_name
colnames(smash_sds_cmp_mat) <- rownames(smash_sds_cmp_mat)

```

#### SDS tree objects
```{r}
# Create Phylo Tree from sourmash matrix

smash_sds_k31_tree_mat <- dist2(smash_sds_cmp_mat)
smash_sds_k31_tree_fit <- hclust(smash_sds_k31_tree_mat)
smash_sds_k31_phylo <- as.phylo(smash_sds_k31_tree_fit)

```


#### SDS ggtree
```{r}
# Make tree
sm_sds_k31_tree <- ggtree(smash_sds_k31_phylo,
                          color = "black",
                          size = 0.4,
                          layout = "fan",
                          open.angle = 15)
```

### Dos-Santos Negative (DSN) - Group C
```{r}
# DSN Group
smash_dsn_cmp <- read.csv("./sourmash_data/dsn_dos-santos-negative/pure_iso_dsn_k31_cmp.csv", header = TRUE, check.names = FALSE)

# set rownames to colnames
rownames(smash_dsn_cmp) <- colnames(smash_dsn_cmp)

# add bcw ids to smash_dsn_cmp data frame
smash_dsn_cmp$bcw_id <- abb_bcw_labels$BCW_ID[match(rownames(smash_dsn_cmp), abb_bcw_labels$BCW_ID)]

# add genus info to smash_dsn_cmp
smash_dsn_cmp$Genus <- smash_lca_whole_genomes$genus[match(smash_dsn_cmp$bcw_id, smash_lca_whole_genomes$bcw_id)]

# add tx_name Variable to smash_dsn_cmp
smash_dsn_cmp$tx_name <- paste(smash_dsn_cmp$genus, sprintf("sp. %s", smash_dsn_cmp$bcw_id))

# add BNF value
smash_dsn_cmp$BNF <- pgp_15N$ratio[match(smash_dsn_cmp$bcw_id, pgp_15N$BCW_ID)]

# set rownames to match sample
rownames(smash_dsn_cmp) <- smash_dsn_cmp$tx_name

# create matrices for tree

## sourmash compare matrix
smash_dsn_cmp_mat <- as.matrix(smash_dsn_cmp[,1:233])
rownames(smash_dsn_cmp_mat) <- smash_dsn_cmp$tx_name
colnames(smash_dsn_cmp_mat) <- rownames(smash_dsn_cmp_mat)

```

#### DSN tree objects
```{r}
# Create Phylo Tree from sourmash matrix

smash_dsn_k31_tree_mat <- dist2(smash_dsn_cmp_mat)
smash_dsn_k31_tree_fit <- hclust(smash_dsn_k31_tree_mat)
smash_dsn_k31_phylo <- as.phylo(smash_dsn_k31_tree_fit)

```


#### DSN ggtree
```{r}
# Make tree
sm_dsn_k31_tree <- ggtree(smash_dsn_k31_phylo,
                          color = "black",
                          size = 0.4,
                          layout = "fan",
                          open.angle = 15)

```

### Cowplot NIF Group ggtrees

> Need to remove geom_tiplab for grid plotting

##### DSP Isolates - no tip
```{r}
# Add gheatmap 0 for genus
## reproducible colors
bnf0_notip_dsp <- gheatmap(sm_dsp_k31_tree, smash_dsp_cmp[, "Genus", drop = F],
                     offset = 0,
                     width = 0.05,
                     colnames_position = "top",
                     colnames_angle = 0,
                     colnames_offset_y = 3.5,
                     font.size = 4.5,
                     color = "grey50") + 
  scale_fill_manual(values=c("#72A890",
                             "#86D6E1",
                             "#77E94F",
                             "#E1ABAA",
                             "#7BEFDF",
                             "#DA47D2",
                             "#CF7DDC"),
                    name = "Genus",
                    guide = guide_legend(
                      direction = "vertical",
                      ncol = 2,
                      title.position = "top",
                      title.theme = element_text(size = 18, hjust = 0.5),
                      label.position = "right",
                      label.theme = element_text(size = 14)
                      )
  ) +
  theme(legend.title = element_text(), legend.key.size = unit(.5, "cm"))

# Add gheatmap 1 for ACC
bnf1_notip_dsp <- bnf0_notip_dsp + new_scale_fill()

bnf2_notip_dsp <- gheatmap(bnf1_notip_dsp, smash_dsp_cmp[, "BNF", drop = F],
                     offset = 0.2,
                     width = 0.05,
                     colnames_position = "top",
                     colnames_angle = 0,
                     colnames_offset_y = 3.5,
                     font.size = 4.5,
                     color = "grey50") + 
  scale_fill_viridis_c(option="inferno",
                       direction = -1,
                       name = "BNF (15N/14N)",
                       limits = c(0,5),
                       breaks = c(0, 1, 2, 3, 4, 5),
                       guide = guide_colorbar(
                         direction = "horizontal",
                         title.position = "top",
                         title.theme = element_text(size = 18, hjust = 0.5),
                         barwidth = 18,
                         label.position = "bottom",
                         label.theme = element_text(size = 14)
                         )) +
  theme(legend.title = element_text(), legend.position = "bottom")

bnf_dsp_rot <- print(rotate_tree(bnf2_notip_dsp, -83))

# save plot
ggsave("./Plots/bnf_dsp_rot.pdf", bnf_dsp_rot, width = 22, height = 14)

```

##### SDS Isolates - no tip
```{r}
# Add gheatmap 0 for genus
## reproducible colors
bnf0_notip_sds <- gheatmap(sm_sds_k31_tree, smash_sds_cmp[, "Genus", drop = F],
                     offset = 0,
                     width = 0.05,
                     colnames_position = "top",
                     colnames_angle = 0,
                     colnames_offset_y = 0.9,
                     font.size = 4.5,
                     color = "grey50") + 
  scale_fill_manual(values=c("#E6BE53",
                             "#CAD773",
                             "#8F3EE4",
                             "#D8E9D6",
                             "#C59D6B",
                             "#64E7A9",
                             "#767CDE",
                             "#E1ABAA",
                             "#B4E3A7",
                             "#CF7DDC"),
                    name = "Genus",
                    guide = guide_legend(
                      direction = "vertical",
                      ncol = 2,
                      title.position = "top",
                      title.theme = element_text(size = 18, hjust = 0.5),
                      label.position = "right",
                      label.theme = element_text(size = 14)
                      )
  ) +
  theme(legend.title = element_text(), legend.key.size = unit(.5, "cm"))

# Add gheatmap 1 for ACC
bnf1_notip_sds <- bnf0_notip_sds + new_scale_fill()

bnf2_notip_sds <- gheatmap(bnf1_notip_sds, smash_sds_cmp[, "BNF", drop = F],
                     offset = 0.125,
                     width = 0.05,
                     colnames_position = "top",
                     colnames_angle = 0,
                     colnames_offset_y = 0.9,
                     font.size = 4.5,
                     color = "grey50") + 
  scale_fill_viridis_c(option="inferno",
                       direction = -1,
                       name = "BNF (15N/14N)",
                       limits = c(0,5),
                       breaks = c(0, 1, 2, 3, 4, 5),
                       guide = guide_colorbar(
                         direction = "horizontal",
                         title.position = "top",
                         title.theme = element_text(size = 18, hjust = 0.5),
                         barwidth = 18,
                         label.position = "bottom",
                         label.theme = element_text(size = 14)
                         )) +
  theme(legend.title = element_text(), legend.position = "bottom")

bnf_sds_rot <- print(rotate_tree(bnf2_notip_sds, -85))

# save plot
ggsave("./Plots/bnf_sds_rot.pdf", bnf_sds_rot, width = 22, height = 14)

```

##### DSN Isolates - no tip
```{r}
# Add gheatmap 0 for genus
## reproducible colors
bnf0_notip_dsn <- gheatmap(sm_dsn_k31_tree, smash_dsn_cmp[, "Genus", drop = F],
                     offset = 0,
                     width = 0.05,
                     colnames_position = "top",
                     colnames_angle = 0,
                     colnames_offset_y = 4.5,
                     font.size = 4.35,
                     color = "grey50") + 
  scale_fill_manual(values=c("#CAD773",
                             "#81D870",
                             "#E25F9D",
                             "#D8E9D6",
                             "#72A890",
                             "#CDEC4C",
                             "#81748E",
                             "#D3D0E3",
                             "#7DAEDE",
                             "#E2684F",
                             "#E1ABAA",
                             "#7BEFDF",
                             "#D7A7DF",
                             "#E5DEA8",
                             "#CF7DDC"),
                    name = "Genus",
                    guide = guide_legend(
                      direction = "vertical",
                      ncol = 2,
                      title.position = "top",
                      title.theme = element_text(size = 18, hjust = 0.5),
                      label.position = "right",
                      label.theme = element_text(size = 14)
                      )
  ) +
  theme(legend.title = element_text(), legend.key.size = unit(.5, "cm"))

# Add gheatmap 1 for BNF
bnf1_notip_dsn <- bnf0_notip_dsn + new_scale_fill()

bnf2_notip_dsn <- gheatmap(bnf1_notip_dsn, smash_dsn_cmp[, "BNF", drop = F],
                     offset = 0.22,
                     width = 0.05,
                     colnames_position = "top",
                     colnames_angle = 0,
                     colnames_offset_y = 4.5,
                     font.size = 4.35,
                     color = "grey50") + 
  scale_fill_viridis_c(option="inferno",
                       direction = -1,
                       name = "BNF (15N/14N)",
                       limits = c(0,5),
                       breaks = c(0, 1, 2, 3, 4, 5),
                       guide = guide_colorbar(
                         direction = "horizontal",
                         title.position = "top",
                         title.theme = element_text(size = 18, hjust = 0.5),
                         barwidth = 18,
                         label.position = "bottom",
                         label.theme = element_text(size = 14)
                         )) +
  theme(legend.title = element_text(), legend.position = "bottom")

bnf_dsn_rot <- print(rotate_tree(bnf2_notip_dsn, -83))

# save plot
ggsave("./Plots/bnf_dsn_rot.pdf", bnf_dsn_rot, width = 22, height = 14)

```

### Extract legends for pgp plots
#### Genus
```{r}
# Genus Tree
cowplot_genus <- gheatmap(sm_k31_tree_o15, pgp_master[, "Genus", drop = F],
                     offset = 0,
                     width = 0.05,
                     colnames_position = "top",
                     colnames_angle = 0,
                     colnames_offset_y = 9.5,
                     font.size = 3.5,
                     color = "grey50") + 
  scale_fill_manual(values=c("#E6BE53",
                             "#CAD773",
                             "#81D870",
                             "#8F3EE4",
                             "#E25F9D",
                             "#D8E9D6",
                             "#72A890",
                             "#CDEC4C",
                             "#C59D6B",
                             "#86D6E1",
                             "#77E94F",
                             "#81748E",
                             "#64E7A9",
                             "#D3D0E3",
                             "#767CDE",
                             "#7DAEDE",
                             "#E2684F",
                             "#E1ABAA",
                             "#7BEFDF",
                             "#DA47D2",
                             "#D7A7DF",
                             "#E5DEA8",
                             "#B4E3A7",
                             "#CF7DDC"),
                    name = "Genus",
                    guide = guide_legend(
                      direction = "vertical",
                      ncol = 3,
                      title.position = "top",
                      title.theme = element_text(size = 18, hjust = 0.5),
                      label.position = "right",
                      label.theme = element_text(size = 16)
                      )) + theme(legend.title = element_text(),
                                 legend.key.size = unit(0.75, "cm"))

cowplot_genus

legend_genus <- get_legend(cowplot_genus)
```


#### BNF
```{r}
cowplot_bnf <- gheatmap(sm_dsp_k31_tree, pgp_master[, "BNF", drop = F],
                     offset = 00,
                     width = 0.05,
                     colnames_position = "top",
                     colnames_angle = 0,
                     colnames_offset_y = 2.25,
                     font.size = 3.5,
                     color = "grey50") + 
  scale_fill_viridis_c(option="inferno",
                       direction = -1,
                       name = "BNF (15N/14N Ratio)",
                       limits = c(0,5),
                       breaks = c(0, 1, 2, 3, 4, 5),
                       guide = guide_colorbar(
                         direction = "horizontal",
                         title.position = "top",
                         title.theme = element_text(size = 18, hjust = 0.5),
                         barwidth = 18,
                         label.position = "bottom",
                         label.theme = element_text(size = 14)
                         )) +
  theme(legend.title = element_text(), legend.position = "bottom")

cowplot_bnf

legend_bnf <- get_legend(cowplot_bnf)

```

### Assay Legend Array
```{r}

bnf_legend_array_1 <- plot_grid(legend_genus,
                              legend_bnf,
                              nrow = 2)

bnf_legend_array_1

```

#### Plot Grid


```{r}
bnf_row <- plot_grid(bnf_dsp_rot + theme(plot.margin = unit(c(0, 0, 0, 0), "cm"),
                                         legend.position = "none"),
                     bnf_sds_rot + theme(plot.margin = unit(c(0, 0, 0, 0), "cm"),
                                         legend.position = "none"),
                     bnf_dsn_rot + theme(plot.margin = unit(c(0, 0, 0, 0), "cm"),
                                         legend.position = "none"),
                     bnf_legend_array_1,
                     labels = c("A", "B", "C", ""),
                     label_size = 24,
                     nrow = 2,
                     ncol = 2,
                     vjust = 3,
                     align = 'vh'
                     )

#pgp_grid <- plot_grid(pgp_row, pgp_all_legend,
#                    ncol = 1,
#                    rel_heights = c(1, 0.5))
bnf_row

ggsave("./plots/Fig4.pdf", bnf_row, width = 22, height = 14)
```

### BNF Assay Summary

#### Min and Max BNF values
```{r}
print("For all single isolate BNF phenotypes,")
print("The min 15N ratio is:")
min(pgp_master[,"BNF"], na.rm = T)
print("The max 15N ratio is:")
max(pgp_master[,"BNF"], na.rm = T)
```

#### NIF Group Summary
```{r}

# DSP Group
## N15 < 1
smash_dsp_cmp %>%
  group_by(Genus, N15) %>%
  select(N15) %>%
  filter(N15 < 1) %>%
  count()

## 1 <= N15 <= 2
smash_dsp_cmp %>%
  group_by(Genus, N15) %>%
  select(N15) %>%
  filter(N15 >= 1 & N15 <=2) %>%
  count()

## N15 >= 2
smash_dsp_cmp %>%
  group_by(Genus, N15) %>%
  select(N15) %>%
  filter(N15 >= 2) %>%
  count()

# SDS Group
## N15 < 1
smash_sds_cmp %>%
  group_by(Genus, N15) %>%
  select(N15) %>%
  filter(N15 < 1) %>%
  count()

## 1 <= N15 <= 2
smash_sds_cmp %>%
  group_by(Genus, N15) %>%
  select(N15) %>%
  filter(N15 >= 1 & N15 <2) %>%
  count()

## N15 >= 2
smash_sds_cmp %>%
  group_by(Genus, N15) %>%
  select(N15) %>%
  filter(N15 >= 2) %>%
  count()

# DSN Group
## N15 < 1
smash_dsn_cmp %>%
  group_by(Genus, N15) %>%
  select(N15) %>%
  filter(N15 < 1) %>%
  count()

## 1 <= N15 <= 2
smash_dsn_cmp %>%
  group_by(Genus, N15) %>%
  select(N15) %>%
  filter(N15 >= 1 & N15 <2) %>%
  count()

## N15 >= 2
smash_dsn_cmp %>%
  group_by(Genus, N15) %>%
  select(N15) %>%
  filter(N15 >= 2) %>%
  count()
```
